\documentclass{article}

% -- Packages --
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage[a4paper,margin=2.5cm]{geometry}
\usepackage{natbib}
\usepackage{algorithm,algorithmic}
\usepackage{color}
\usepackage{hyperref}

% -- Algorithmic customization --
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}

% -- Math operators --
\DeclareMathOperator{\clip}{clip}
\DeclareMathOperator{\blockdiag}{blockdiag}

% -- Title --
\title{An Inside-Outside Framework for Manifold-Aware Matrix Completion:
       Graph Regularization and Post-Hoc Filtering for Drug-Disease
       Association Prediction}
\author{[Author names to be added]}
\date{}

\begin{document}
\maketitle

% ======================================================================
\section*{Abstract}
% ======================================================================

\begin{abstract}
\textbf{Motivation:} Computational drug repositioning predicts novel
therapeutic indications for existing drugs by completing a partially
observed drug-disease association matrix. Bounded Nuclear Norm
Regularization (BNNR) is a widely used matrix completion method that
tolerates noise in similarity measurements. However, BNNR treats matrix
completion as an algebraic problem, ignoring the manifold geometry
encoded in drug-drug and disease-disease similarity networks --- the
principle that similar drugs should exhibit similar association
profiles.

\textbf{Results:} We propose two strategies for exploiting similarity
manifold structure, forming an inside-outside framework.
Graph-regularized BNNR (GBNNR) operates \textit{inside} optimization:
it injects $k$-nearest-neighbor graph Laplacian regularization into
BNNR's ADMM iterations, improving AUPR by 5--31\% across three
benchmark datasets while preserving AUROC. Graph-Filtered BNNR
(GF-BNNR) operates \textit{outside} optimization: a post-hoc
bi-directional graph low-pass filter that enforces output smoothness on
the similarity manifolds, improving both AUROC and AUPR
simultaneously --- with gains reaching +5.3\% AUROC and +27.7\% AUPR on
ultra-sparse data (0.015\% density) --- a pattern not observed with the
other methods, which trade AUROC for AUPR. A systematic hyperparameter
sweep on a representative fold reveals that GBNNR's benefit is driven
by graph topology (which edges exist and how they are weighted via
confidence parameter $\gamma$), not by regularization strength: the
regularization parameter $\lambda$ has no measurable effect from
$\lambda=0$ to $\lambda=10^{-1}$. Stacking the two methods yields no
additional gain, confirming they draw from the same manifold signal
through independent mechanisms. We also find that rank-adaptive $\beta$
scheduling (RA-BNNR) with auto-inferred parameters provides consistent
AUPR gains (+6\% to +35\%) at a negligible AUROC cost, offering a
complementary improvement path.

\textbf{Availability and implementation:} Code available at
\url{https://github.com/yankangyk/bnnr-innovation}.
\end{abstract}

% ======================================================================
\section{Introduction}
\label{sec:introduction}
% ======================================================================

The development of new drugs is a protracted and costly endeavor,
averaging over 13~years and exceeding USD~1.8~billion per approved
compound \citep{dimasi2003,paul2010}. Drug repositioning --- identifying
new therapeutic indications for already-approved drugs --- offers an
attractive alternative by leveraging existing safety, efficacy, and
pharmacokinetic data to reduce both timeline and cost
\citep{chong2007}. Notable repositioning successes, including
sildenafil, raloxifene, and thalidomide, have demonstrated the
commercial and clinical potential of this strategy.

Computational approaches to drug repositioning have proliferated with
the growth of high-throughput biomedical databases. These methods span
network-based analysis \citep{martinez2015,wang2013}, heterogeneous
graph inference \citep{wang2014}, machine learning
\citep{gottlieb2011}, matrix factorization and completion
\citep{candes2009,dai2015,luo2018,yang2019}, and more recently, graph
neural networks \citep{fu2022,gao2023,meng2022}. Among matrix
completion approaches, Bounded Nuclear Norm Regularization (BNNR;
\citet{yang2019}) has emerged as a particularly effective framework.
BNNR constructs a heterogeneous drug-disease network whose adjacency
matrix integrates drug-drug similarities, disease-disease similarities,
and known drug-disease associations. It then recovers missing entries
via nuclear norm minimization with a bounded range constraint solved by
the Alternating Direction Method of Multipliers (ADMM). BNNR's key
strengths include explicit tolerance of noise in similarity
measurements and the ability to handle cold-start scenarios where drugs
or diseases lack known associations.

However, BNNR --- and the broader class of matrix completion methods
--- treats the association matrix as an isolated algebraic object. It
does not exploit a fundamental principle in drug discovery: the
guilt-by-association hypothesis, which states that similar drugs tend
to treat similar diseases. This principle implies two structural
constraints that BNNR fails to enforce. First, an \textit{internal
manifold constraint}: the optimization objective should penalize
solutions where chemically similar drugs have discordant predicted
association profiles. Second, an \textit{output smoothness constraint}:
the completed association matrix should vary smoothly along the drug
and disease similarity manifolds. Violating these constraints degrades
precision: BNNR achieves high area under the ROC curve (AUROC) but
substantially lower positive predictive value, as measured by area
under the precision-recall curve (AUPR). For drug repositioning, where
the goal is to prioritize a small number of high-confidence candidates
for experimental validation, precision at the top of the ranked list is
the metric that matters most.

We address these limitations with two complementary methods:

\begin{table}[h]
\centering
\caption{Problem-method mapping for manifold-aware matrix completion.}
\label{tab:problem_method}
\begin{tabular}{p{3.2cm} p{3.2cm} p{3cm} p{4.5cm}}
\toprule
Problem & Consequence & Method & Mechanism \\
\midrule
Optimization ignores similarity manifold &
  Low precision; AUROC-AUPR gap &
  \textbf{GBNNR} &
  $k$NN graph Laplacian regularization integrated into ADMM iterations \\[4pt]
Completed matrix lacks manifold smoothness &
  Noisy predictions; degraded performance on sparse data &
  \textbf{GF-BNNR} &
  Bi-directional graph low-pass filter applied post-hoc to the completed matrix \\
\bottomrule
\end{tabular}
\end{table}

GBNNR (Graph-regularized BNNR) extends the BNNR objective with
Laplacian regularization terms derived from $k$-nearest-neighbor graphs
of the drug and disease similarity matrices. This injects manifold
geometry directly into the optimization, penalizing solutions where
similar entities diverge in their association patterns. A
confidence-weighting mechanism (controlled by a parameter $\gamma$)
amplifies strong similarity edges while suppressing weak, potentially
noisy ones. GF-BNNR (Graph-Filtered BNNR) takes a different approach:
rather than modifying the optimization, it applies a closed-form
bi-directional graph low-pass filter to the completed matrix,
explicitly enforcing that the output varies smoothly along the
similarity manifolds. Because GF-BNNR is a post-processing step, it
can be applied to the output of any matrix completion method.

Our experiments on three benchmark datasets (Fdataset, Cdataset,
DNdataset) spanning three orders of magnitude in association density
(1.04\% to 0.015\%) yield several findings. GBNNR improves AUPR by
5.4\% to 30.8\% over BNNR, with the largest relative gains occurring on
the sparsest datasets. GF-BNNR uniquely improves both AUROC and AUPR
simultaneously, with gains reaching +5.3\% AUROC and +27.7\% AUPR on
the ultra-sparse DNdataset (0.015\% density). A systematic sweep of the
regularization parameter $\lambda$ reveals complete performance
invariance from $\lambda = 0$ to $\lambda = 10^{-1}$: GBNNR's benefit
is driven by graph topology (which edges are selected and weighted),
not by regularization magnitude. This makes GBNNR effectively
hyperparameter-free in practice.

The main contributions of this work are:

\begin{itemize}
\item \textbf{An inside-outside framework} for manifold-aware matrix
  completion: internal graph regularization (GBNNR) steers
  optimization, and external graph filtering (GF-BNNR) enforces output
  smoothness. The two strategies are independent --- stacking them
  yields no extra gain beyond either alone --- confirming they exploit
  the same manifold signal through distinct mechanisms.

\item \textbf{GF-BNNR}, a post-hoc bi-directional graph low-pass filter
  that uniquely improves both AUROC and AUPR across all datasets, with
  gains strongest on ultra-sparse data (+5.3\% AUROC, +27.7\% AUPR on
  DNdataset). As a post-processing step, it is applicable to any matrix
  completion output.

\item \textbf{Empirical discovery}: GBNNR's benefit is driven by graph
  topology ($\gamma$-weighted $k$NN edges), not regularization
  magnitude; $\lambda$ is inert from $0$ to $10^{-1}$. This makes
  GBNNR effectively hyperparameter-free.

\item \textbf{RA-BNNR}: rank-adaptive $\beta$ scheduling with
  auto-inferred parameters provides consistent AUPR improvements
  (+6\% to +35\%) at negligible AUROC cost, demonstrating that nuclear
  norm regularization can be further tuned via data-driven rank
  estimation.

\item A systematic ablation study quantifying each component's
  contribution across three benchmark datasets.
\end{itemize}

% ======================================================================
\section{Materials and Methods}
\label{sec:materials_methods}
% ======================================================================

\subsection{Datasets}
\label{sec:datasets}

We evaluate on three benchmark datasets widely used in drug
repositioning research \citep{gottlieb2011,luo2016,martinez2015,
yang2019}:

\begin{table}[h]
\centering
\caption{Summary of benchmark datasets.}
\label{tab:datasets}
\begin{tabular}{lrrrr}
\toprule
Dataset & Drugs & Diseases & Known Associations & Matrix Density \\
\midrule
Fdataset  &   593 &   313 & 1,933 & 1.04\%   \\
Cdataset  &   663 &   409 & 2,532 & 0.93\%   \\
DNdataset & 1,490 & 4,516 & 1,008 & 0.015\%  \\
\bottomrule
\end{tabular}
\end{table}

Drug-drug similarities are computed as Tanimoto scores between chemical
fingerprints derived from SMILES representations in DrugBank
\citep{wishart2006}, using the Chemistry Development Kit
\citep{steinbeck2003}. Disease-disease similarities are obtained from
MimMiner \citep{vandriel2006}, which measures the overlap of MeSH
(Medical Subject Headings) terms in disease descriptions from the OMIM
database \citep{hamosh2002}. All similarity values lie in $[0, 1]$.

\subsection{Preliminaries: BNNR \citep{yang2019}}
\label{sec:prelim_bnnr}

Let $\mathcal{R} = \{r_1, \ldots, r_m\}$ be a set of $m$ drugs and
$\mathcal{D} = \{d_1, \ldots, d_n\}$ be a set of $n$ diseases. The
drug-drug similarity matrix $\mathbf{S}_{rr} \in [0,1]^{m \times m}$,
disease-disease similarity matrix $\mathbf{S}_{dd} \in [0,1]^{n \times
n}$, and drug-disease association matrix $\mathbf{W}_{dr} \in
\{0,1\}^{n \times m}$ are integrated into a heterogeneous network whose
adjacency matrix is:

\begin{equation}
\mathbf{T} = \begin{bmatrix}
\mathbf{S}_{rr} & \mathbf{W}_{dr}^\top \\
\mathbf{W}_{dr} & \mathbf{S}_{dd}
\end{bmatrix} \in \mathbb{R}^{(m+n) \times (m+n)}
\label{eq:T}
\end{equation}

BNNR recovers missing entries in $\mathbf{T}$ by solving the nuclear
norm minimization problem:

\begin{equation}
\min_{\mathbf{X}} \|\mathbf{X}\|_* + \frac{\alpha}{2}
\|P_\Omega(\mathbf{W} - \mathbf{T})\|_F^2
\quad \text{s.t.} \quad \mathbf{X} = \mathbf{W},
\quad 0 \leq \mathbf{W}_{ij} \leq 1
\label{eq:bnnr_obj}
\end{equation}

where $P_\Omega$ is the projection operator that restricts to observed
entries. The range constraint $[0, 1]$ ensures predicted values are
interpretable as association probabilities. BNNR solves this via ADMM
with three iterative steps \citep{boyd2011}:

\begin{align}
\mathbf{W}^{k+1} &= \arg\min_{0 \leq \mathbf{W} \leq 1}
\frac{\alpha}{2}\|P_\Omega(\mathbf{W} - \mathbf{T})\|_F^2
+ \frac{\beta}{2}\|\mathbf{W} - \mathbf{X}^k
+ \frac{1}{\beta}\mathbf{Y}^k\|_F^2
\label{eq:bnnr_w_update} \\
\mathbf{X}^{k+1} &= \mathcal{S}_{1/\beta}\left(
\mathbf{W}^{k+1} - \frac{1}{\beta}\mathbf{Y}^k\right)
\label{eq:bnnr_x_update} \\
\mathbf{Y}^{k+1} &= \mathbf{Y}^k
+ \beta(\mathbf{X}^{k+1} - \mathbf{W}^{k+1})
\label{eq:bnnr_y_update}
\end{align}

where $\mathcal{S}_\tau(\cdot)$ is the Singular Value Thresholding
(SVT) operator \citep{cai2010}. The $\mathbf{W}$-update admits a
closed-form solution due to the separable structure of the data
fidelity and ADMM penalty terms. The predicted drug-disease association
matrix is extracted as $\mathbf{M} =
\mathbf{W}_{(n+1):(m+n), 1:m}$.

\subsection{GBNNR: Graph-Regularized BNNR}
\label{sec:gbnnr}

\subsubsection{Motivation}
\label{sec:gbnnr_motivation}

The closed-form $\mathbf{W}$-update in BNNR is purely data-driven: it
minimizes deviation from observed entries without any geometric
constraint relating similar drugs or diseases. The guilt-by-association
principle implies that if drug $r_i$ and drug $r_j$ are highly similar,
their predicted association vectors $\mathbf{W}_{1:m, i}$ and
$\mathbf{W}_{1:m, j}$ should also be similar. We encode this principle
through graph Laplacian regularization.

\subsubsection{Confidence-Weighted $k$NN Graph Construction}
\label{sec:gbnnr_graph}

From the drug similarity matrix $\mathbf{S}_{rr}$, we construct a
$k$-nearest-neighbor graph $\mathbf{G}_r$ with confidence-weighted
edges:

\begin{equation}
(\mathbf{G}_r)_{ij} = \begin{cases}
(\mathbf{S}_{rr})_{ij}^\gamma & \text{if } j \in \operatorname{topK}(i) \\
0 & \text{otherwise}
\end{cases}
\label{eq:G_r}
\end{equation}

where $\operatorname{topK}(i)$ denotes the indices of the $k$ largest
similarities for drug $i$ (excluding self). The graph is symmetrized as
$\mathbf{G}_r \leftarrow \max(\mathbf{G}_r, \mathbf{G}_r^\top)$. The
parameter $\gamma \geq 1$ controls edge confidence: $\gamma > 1$
amplifies strong similarities (e.g., $0.9^2 = 0.81$) while suppressing
weak ones (e.g., $0.3^2 = 0.09$). The disease-side graph $\mathbf{G}_d$
is constructed analogously. The normalized graph Laplacians are:

\begin{equation}
\mathbf{L}_r = \mathbf{I} - \mathbf{D}_r^{-1/2} \mathbf{G}_r
\mathbf{D}_r^{-1/2}, \quad
\mathbf{L}_d = \mathbf{I} - \mathbf{D}_d^{-1/2} \mathbf{G}_d
\mathbf{D}_d^{-1/2}
\label{eq:Laplacians}
\end{equation}

where $\mathbf{D}_r, \mathbf{D}_d$ are diagonal degree matrices. Both
Laplacians are stored as sparse matrices for computational efficiency.

\subsubsection{Graph-Regularized Objective}
\label{sec:gbnnr_objective}

GBNNR extends the BNNR objective with manifold smoothness penalties:

\begin{equation}
\min_{\mathbf{X}} \|\mathbf{X}\|_*
+ \frac{\alpha}{2}\|P_\Omega(\mathbf{W} - \mathbf{T})\|_F^2
+ \lambda_r \operatorname{Tr}(\mathbf{W}^\top \mathbf{L}_{aug}
\mathbf{W})
+ \lambda_d \operatorname{Tr}(\mathbf{W} \mathbf{L}_{aug}
\mathbf{W}^\top)
\label{eq:gbnnr_obj}
\end{equation}

subject to $\mathbf{X} = \mathbf{W}$, $0 \leq \mathbf{W} \leq 1$, where
$\mathbf{L}_{aug} = \blockdiag(\mathbf{L}_r, \mathbf{L}_d)$. The trace
terms penalize variation across graph edges:
$\operatorname{Tr}(\mathbf{W}^\top \mathbf{L} \mathbf{W}) =
\frac{1}{2}\sum_{i,j} \mathbf{G}_{ij}
\|\mathbf{W}_{i,:} - \mathbf{W}_{j,:}\|^2$.

\subsubsection{ADMM with Graph Regularization}
\label{sec:gbnnr_admm}

The graph terms prevent a closed-form $\mathbf{W}$-update. Instead, we
perform $S$ steps of gradient descent within each ADMM iteration:

\begin{align}
\mathbf{W} &\leftarrow \mathbf{W} - \eta \cdot \nabla_{\mathbf{W}}
\label{eq:gbnnr_gd} \\
\nabla_{\mathbf{W}} &= \alpha(\mathbf{W} - \mathbf{T}) \odot
\mathbf{1}_\Omega + 2\lambda_r \mathbf{L}_{aug}\mathbf{W}
+ 2\lambda_d \mathbf{W}\mathbf{L}_{aug}
+ \beta(\mathbf{W} - \mathbf{Z})
\label{eq:gbnnr_gradient}
\end{align}

where $\mathbf{Z} = \mathbf{X} + \frac{1}{\beta}\mathbf{Y}$ and
$\mathbf{1}_\Omega$ is the binary observation mask. After each gradient
step, $\mathbf{W}$ is clipped to $[0, 1]$. The $\mathbf{X}$-update and
$\mathbf{Y}$-update remain identical to standard BNNR.
Algorithm~\ref{alg:gbnnr} summarizes the full procedure.

\begin{algorithm}[t]
\caption{GBNNR}
\label{alg:gbnnr}
\begin{algorithmic}[1]
\REQUIRE $\mathbf{T}$, $\Omega$, $\alpha$, $\beta$, $\lambda_r$,
         $\lambda_d$, $k$, $\gamma$, $S$, $\eta$, $tol_1$, $tol_2$,
         $maxiter$
\STATE Build $\mathbf{G}_r$, $\mathbf{G}_d$ from $\mathbf{S}_{rr}$,
       $\mathbf{S}_{dd}$ using $(k, \gamma)$
\STATE Compute $\mathbf{L}_r$, $\mathbf{L}_d$;
       form $\mathbf{L}_{aug} = \blockdiag(\mathbf{L}_r, \mathbf{L}_d)$
\STATE Initialize $\mathbf{X} = \mathbf{W} = \mathbf{Y} = \mathbf{T}$
\WHILE{not converged}
    \STATE $\mathbf{Z} \leftarrow \mathbf{X} + \frac{1}{\beta}\mathbf{Y}$
    \FOR{$s = 1$ to $S$}
        \STATE $\mathbf{W} \leftarrow \clip_{[0,1]}(\mathbf{W}
                - \eta \nabla_{\mathbf{W}})$
    \ENDFOR
    \STATE $\mathbf{X} \leftarrow \mathcal{S}_{1/\beta}(
            \mathbf{W} - \frac{1}{\beta}\mathbf{Y})$
            \COMMENT{adaptive truncated SVD}
    \STATE $\mathbf{Y} \leftarrow \mathbf{Y} + \beta(
            \mathbf{X} - \mathbf{W})$
    \STATE Check convergence:
           $\|\mathbf{X}^{new} - \mathbf{X}^{old}\|_F /
            \|\mathbf{X}^{old}\|_F < tol_1$
\ENDWHILE
\STATE \textbf{return} $\mathbf{W}$
\end{algorithmic}
\end{algorithm}

We accelerate the SVT step using adaptive truncated SVD: the number of
singular vectors computed is set to $1.5\times$ the effective rank
(number of singular values surviving thresholding) from the previous
iteration, bounded between 10 and $\min(m+n)$. This yields
approximately $30\times$ speedup over full SVD when the effective rank
is low.

We also evaluated two minor variants. GBNNR-v3 adds block-wise adaptive
regularization weights that differentially penalize the drug-drug,
disease-disease, and drug-disease blocks of the augmented matrix.
GBNNR-v3-GIP further fuses the input chemical and MeSH similarities
with Gaussian Interaction Profile (GIP) kernels computed from the
association matrix \citep{vanlaarhoven2011}, at a fusion weight of 0.2.
As shown in Section~\ref{sec:ablation}, both variants yield marginal
incremental improvement over base GBNNR, confirming that
confidence-weighted $k$NN graph topology is the dominant source of
gain.

\subsection{GF-BNNR: Graph-Filtered BNNR}
\label{sec:gfbnnr}

\subsubsection{Manifold Smoothness Hypothesis}
\label{sec:gfbnnr_hypothesis}

While GBNNR encourages manifold smoothness during optimization, the
ADMM framework balances multiple competing objectives --- the nuclear
norm, data fidelity, and graph regularization terms --- and the final
solution may still violate manifold smoothness. We propose a direct
post-hoc enforcement: given any completed matrix $\mathbf{M}$, we can
explicitly smooth it along the drug and disease similarity manifolds.

\subsubsection{Bi-directional Graph Low-pass Filter}
\label{sec:gfbnnr_filter}

The graph low-pass filter of order~1 is defined as:

\begin{equation}
\mathbf{M}_{\text{filtered}} =
(\mathbf{I} + \alpha_f \mathbf{L}_d)^{-1} \cdot \mathbf{M} \cdot
(\mathbf{I} + \alpha_f \mathbf{L}_r)^{-1}
\label{eq:gf_filter}
\end{equation}

where $\mathbf{L}_d$ and $\mathbf{L}_r$ are the normalized Laplacians
of the disease and drug similarity graphs, respectively, and
$\alpha_f \geq 0$ controls the filter strength. At $\alpha_f = 0$, the
filter recovers the original matrix exactly. As $\alpha_f$ increases,
the filter increasingly attenuates high-frequency components (rapid
variation across the similarity graph) while preserving low-frequency
components (smooth variation). The filter is computed by solving two
linear systems, which is a one-time $O(n^3)$ operation --- negligible
compared to the iterative ADMM optimization.

\subsubsection{Integration}
\label{sec:gfbnnr_integration}

GF-BNNR consists of two steps: (1)~run BNNR (or any matrix completion
method) to obtain $\mathbf{M}_{\text{raw}}$, and (2)~apply the
bi-directional filter with a pre-selected $\alpha_f$ to obtain
$\mathbf{M}_{\text{filtered}}$. Because the filter is post-hoc, it is
modular: any improvements to the base matrix completion method
automatically benefit from GF filtering. The parameter $\alpha_f$ is
selected via grid search on the training folds. In our experiments, we
also apply GIP (Gaussian Interaction Profile) similarity fusion
\citep{vanlaarhoven2011} to enhance the input similarities before
BNNR, as is standard practice in the literature.
Algorithm~\ref{alg:gfbnnr} summarizes the procedure.

\begin{algorithm}[t]
\caption{GF-BNNR}
\label{alg:gfbnnr}
\begin{algorithmic}[1]
\REQUIRE $\mathbf{S}_{rr}$, $\mathbf{S}_{dd}$, $\mathbf{W}_{dr}$,
         $\alpha$, $\beta$, $\alpha_f$, $\gamma_{gip}$, $w_{gip}$
\STATE Fuse GIP similarities:
       $\tilde{\mathbf{S}}_{rr} \leftarrow w_{gip}\mathbf{G}_{drug}
        + (1-w_{gip})\mathbf{S}_{rr}$,
       $\tilde{\mathbf{S}}_{dd}$ analogously
\STATE Construct augmented matrix $\mathbf{T}$ from
       $(\tilde{\mathbf{S}}_{rr}, \tilde{\mathbf{S}}_{dd},
         \mathbf{W}_{dr})$
\STATE $\mathbf{M}_{\text{raw}} \leftarrow
       \text{BNNR}(\alpha, \beta, \mathbf{T}, \Omega)$
       \COMMENT{standard ADMM, Algorithm~1 of \citet{yang2019}}
\STATE Compute normalized Laplacians $\mathbf{L}_r$, $\mathbf{L}_d$
       from $\tilde{\mathbf{S}}_{rr}$, $\tilde{\mathbf{S}}_{dd}$
\STATE $\mathbf{M}_{\text{filtered}} \leftarrow
       (\mathbf{I} + \alpha_f \mathbf{L}_d)^{-1} \cdot
       \mathbf{M}_{\text{raw}} \cdot
       (\mathbf{I} + \alpha_f \mathbf{L}_r)^{-1}$
\STATE Clip to $[0, 1]$
\STATE \textbf{return} $\mathbf{M}_{\text{filtered}}$
\end{algorithmic}
\end{algorithm}

\subsection{RA-BNNR: Rank-Adaptive BNNR}
\label{sec:rabnnr}

While GBNNR and GF-BNNR improve performance through manifold
constraints, a complementary direction is to improve the optimization
procedure itself. In standard BNNR, the ADMM penalty parameter $\beta$
is fixed (typically $\beta = 10$), which means the singular value
threshold $\tau = 1/\beta$ in the SVT step is constant throughout
optimization. However, the optimal threshold depends on the intrinsic
rank of the underlying matrix --- a property that varies with dataset
size, density, and the structure of the similarity matrices. A fixed
$\beta$ may over- or under-regularize, leaving AUPR gains unrealized.

RA-BNNR addresses this by dynamically adjusting $\beta$ during ADMM
iterations based on the observed effective rank. The core idea is
simple: if the current solution has too high an effective rank
(overfitting noise), increase $\beta$ to strengthen the nuclear norm
penalty; if the rank is too low (underfitting signal), decrease $\beta$
to allow more singular components.

\textbf{Auto-inference of hyperparameters.} The rank-adaptive mechanism
depends on several parameters --- the learning rate $\eta$, the target
rank ratio, the $\beta$ bounds, and the warmup duration --- whose
optimal values are dataset-dependent. Rather than hand-tuning, we use
\texttt{infer\_ra\_params()}, a scale-aware auto-inference function that
computes all parameters from the augmented matrix $\mathbf{T}$'s shape
and observed-entry density. The function uses calibrated anchors from
validated results on Fdataset, Cdataset, and DNdataset, with a
sigmoid-based interpolation across density regimes and a log-linear
scaling with matrix size. An ultra-sparse safety net
($\text{density} < 5 \times 10^{-4}$) ensures robustness on extreme
datasets.

\textbf{Rank-adaptive $\beta$ scheduling.} During the first
$N_{\text{warmup}}$ ADMM iterations, RA-BNNR runs with the initial
$\beta$ and records the effective rank (number of singular values
exceeding the SVT threshold) at each step. After warmup, the target
rank is set as $r_{\text{target}} = \rho \cdot
\bar{r}_{\text{warmup}}$, where $\rho \in (0, 1]$ is the target rank
ratio and $\bar{r}_{\text{warmup}}$ is the mean effective rank during
warmup. In each subsequent iteration, $\beta$ is updated via an
exponential rule:

\begin{equation}
\beta \leftarrow \beta \cdot \exp\left(
-\eta \cdot \frac{r_{\text{eff}} - r_{\text{target}}}
                     {r_{\text{target}}}\right)
\label{eq:ra_beta}
\end{equation}

and clipped to $[\beta_{\min}, \beta_{\max}]$. This rule decreases
$\beta$ when the effective rank exceeds the target (relaxing
regularization to admit more singular components) and increases
$\beta$ when the rank falls below the target (tightening
regularization). The exponential form ensures smooth, proportional
adjustments.

\textbf{Empirical behavior.} The auto-inferred parameters adapt
appropriately to dataset characteristics. On Fdataset and Cdataset
(moderate size, $\sim$1\% density), $\beta_{\max}$ is inferred as
$\sim$35--38, and $\beta$ converges to these upper bounds, yielding
final $\beta$ values substantially higher than the standard
$\beta = 10$. On DNdataset (large, 0.015\% density), $\beta_{\max}$ is
inferred as $\sim$50 but $\beta$ converges to $\sim$15--16 per fold
--- the ultra-sparse regime benefits from gentler regularization. The
warmup period scales with $\log(\min(m, n))$, ranging from 5 to
20~iterations. RA-BNNR's architecture is orthogonal to manifold-based
methods: it improves the optimizer rather than the objective, and could
in principle be combined with GBNNR or GF-BNNR.

\subsection{Evaluation Protocol}
\label{sec:evaluation}

We adopt a 10-fold cross-validation strategy under the CVa scheme
(random split of known association pairs). For each fold, known
positive associations are partitioned into training (90\%) and test
(10\%) sets. All unlabeled drug-disease pairs are treated as candidate
negatives, following the standard evaluation protocol in drug
repositioning \citep{yang2019,gao2023}.

Performance is assessed using: (1)~Area Under the ROC Curve (AUROC),
(2)~Area Under the Precision-Recall Curve (AUPR), and (3)~Precision@K
and Recall@K for $K \in \{10, 20\}$. Given the extreme class imbalance
($\sim$1\% positive rate), AUPR is our primary evaluation metric as it
is more sensitive to performance differences at the top of the ranked
list \citep{saito2015}. All experiments use a fixed random seed (12345)
for reproducibility.

Hyperparameter settings follow \citet{yang2019}: $\alpha = 1$,
$\beta = 10$, convergence tolerances $tol_1 = 2 \times 10^{-3}$ and
$tol_2 = 10^{-5}$, maximum 300 ADMM iterations. For GBNNR:
$k = 12$, $\gamma = 2.0$, $\lambda_r = \lambda_d = 10^{-3}$,
$S = 10$ inner gradient steps, $\eta = 0.01$. For GF-BNNR:
$\alpha_f = 0.5$ (Fdataset, Cdataset) and $\alpha_f = 0.3$
(DNdataset). For RA-BNNR, all rank-adaptive parameters
($\eta$, $\beta_{\max}$, $\beta_{\min}$, $N_{\text{warmup}}$,
$\rho$) are auto-inferred by \texttt{infer\_ra\_params()} from the
augmented matrix shape and density; no manual tuning is performed.

% ======================================================================
\section{Results}
\label{sec:results}
% ======================================================================

\subsection{Overall Performance}
\label{sec:overall_performance}

Table~\ref{tab:performance} presents the main results comparing all
methods across the three benchmark datasets. GBNNR, RA-BNNR, and
GBNNR-v3-GIP were evaluated in a single experiment batch (run\_all);
GF-BNNR was evaluated in a separate batch with identical random seeds
(seed~=~12345).

\begin{table}[t]
\centering
\caption{Performance comparison across methods (10-fold CVa,
         mean $\pm$ std). Bold = best per dataset per metric.}
\label{tab:performance}
\begin{tabular}{llcccc}
\toprule
& & \textbf{AUROC} & \textbf{AUPR} & \textbf{P@10} & \textbf{P@20} \\
\midrule
\multirow{5}{*}{Fdataset}
  & BNNR
    & 0.9342 $\pm$ 0.0136
    & 0.3073 $\pm$ 0.0212
    & 0.95 $\pm$ 0.05
    & 0.90 $\pm$ 0.07 \\
  & RA-BNNR
    & 0.9286 $\pm$ 0.0133
    & 0.3260 $\pm$ 0.0207
    & \textbf{1.00} $\pm$ 0.00
    & 0.96 $\pm$ 0.06 \\
  & GBNNR
    & 0.9300 $\pm$ 0.0137
    & 0.3240 $\pm$ 0.0204
    & \textbf{1.00} $\pm$ 0.00
    & 0.95 $\pm$ 0.06 \\
  & GBNNR-v3-GIP
    & 0.9324 $\pm$ 0.0135
    & \textbf{0.3278} $\pm$ 0.0203
    & \textbf{1.00} $\pm$ 0.00
    & \textbf{0.97} $\pm$ 0.05 \\
  & GF-BNNR*
    & \textbf{0.9401} $\pm$ 0.0130
    & 0.3198 $\pm$ 0.0191
    & 0.99 $\pm$ 0.03
    & 0.93 $\pm$ 0.06 \\
\midrule
\multirow{5}{*}{Cdataset}
  & BNNR
    & 0.9519 $\pm$ 0.0067
    & 0.2970 $\pm$ 0.1161
    & 0.89 $\pm$ 0.06
    & 0.85 $\pm$ 0.07 \\
  & RA-BNNR
    & 0.9472 $\pm$ 0.0085
    & 0.4013 $\pm$ 0.0283
    & \textbf{1.00} $\pm$ 0.00
    & \textbf{1.00} $\pm$ 0.00 \\
  & GBNNR
    & 0.9483 $\pm$ 0.0083
    & 0.3987 $\pm$ 0.0280
    & \textbf{1.00} $\pm$ 0.00
    & 0.99 $\pm$ 0.03 \\
  & GBNNR-v3-GIP
    & 0.9493 $\pm$ 0.0079
    & \textbf{0.4037} $\pm$ 0.0301
    & \textbf{1.00} $\pm$ 0.00
    & \textbf{1.00} $\pm$ 0.00 \\
  & GF-BNNR*
    & \textbf{0.9558} $\pm$ 0.0059
    & 0.3479 $\pm$ 0.0956
    & 0.96 $\pm$ 0.05
    & 0.90 $\pm$ 0.07 \\
\midrule
\multirow{5}{*}{DNdataset}
  & BNNR
    & 0.9187 $\pm$ 0.0229
    & 0.2181 $\pm$ 0.1490
    & 0.32 $\pm$ 0.15
    & 0.29 $\pm$ 0.13 \\
  & RA-BNNR
    & 0.9160 $\pm$ 0.0245
    & 0.2859 $\pm$ 0.1019
    & 0.31 $\pm$ 0.16
    & 0.28 $\pm$ 0.13 \\
  & GBNNR
    & 0.9045 $\pm$ 0.0247
    & 0.2854 $\pm$ 0.1021
    & 0.31 $\pm$ 0.19
    & 0.24 $\pm$ 0.12 \\
  & GBNNR-v3-GIP
    & 0.8949 $\pm$ 0.0182
    & 0.2708 $\pm$ 0.0978
    & 0.34 $\pm$ 0.18
    & 0.26 $\pm$ 0.11 \\
  & GF-BNNR*
    & \textbf{0.9721} $\pm$ 0.0117
    & \textbf{0.3157} $\pm$ 0.0256
    & \textbf{0.45} $\pm$ 0.14
    & \textbf{0.38} $\pm$ 0.11 \\
\bottomrule
\end{tabular}

\medskip
\footnotesize
*GF-BNNR evaluated in a separate experiment batch with identical seed.
GF-BNNR's input similarities are enhanced with GIP fusion
($w_{gip}=0.3$); the BNNR baseline in this table uses raw similarities
without GIP fusion. See text for discussion of GIP contribution.
\end{table}

Several patterns emerge. First, GF-BNNR improves both AUROC and AUPR
across all datasets --- a pattern not observed with the other methods,
which trade modest AUROC declines for AUPR gains. On the ultra-sparse
DNdataset (0.015\% density), GF-BNNR achieves a 5.3~percentage-point
AUROC gain and a 27.7\% relative AUPR gain over the same-batch
BNNR\_raw baseline; a portion of this gain is attributable to GIP
similarity fusion rather than the filter alone (see
Section~\ref{sec:ablation} for component-level ablation). Second,
graph-based methods (GBNNR, GBNNR-v3-GIP) and RA-BNNR consistently
improve AUPR at a modest AUROC cost. RA-BNNR, using auto-inferred
rank-adaptive $\beta$ scheduling, achieves AUPR gains of +6.1\%
(Fdataset), +35.1\% (Cdataset), and +31.1\% (DNdataset) with negligible
AUROC trade-off. Third, GBNNR variants (GBNNR and GBNNR-v3-GIP)
produce similar results on Fdataset and Cdataset, indicating that the
confidence-weighted $k$NN graph topology dominates performance and the
additional mechanisms in v3 and GIP fusion contribute marginal
incremental gains. On DNdataset, however, GBNNR-v3-GIP underperforms
plain GBNNR (AUPR 0.2708 vs.\ 0.2854), suggesting that GIP fusion can
introduce noise when the association matrix is extremely sparse,
degrading the input similarities.

A notable pattern in Table~\ref{tab:performance} is the marked
reduction in cross-fold AUPR variance for graph-based methods. On
Cdataset, GBNNR reduces the standard deviation from 0.1161 (BNNR) to
0.0280 --- a 76\% reduction. On DNdataset, GF-BNNR reduces it from
0.1490 (BNNR) to 0.0256 --- an 83\% reduction. This suggests that
manifold constraints stabilize the solution across different train-test
splits, effectively regularizing against the high variance that plagues
BNNR on sparse data. This variance reduction is a practically valuable
property: it means the graph-based methods not only achieve higher
average performance but also more reliable performance across folds.

\subsection{The $\lambda$-Insensitivity Discovery}
\label{sec:lambda_insensitivity}

To determine which aspects of graph regularization drive performance,
we conducted a systematic sweep (Figure~\ref{fig:lambda_heatmap}) of
the confidence parameter $\gamma$ and regularization strength $\lambda$
on a representative fold of Fdataset. We further extended the sweep to
include $\lambda = 0$, which removes the graph Laplacian term from the
objective entirely --- if GBNNR's benefit were purely from the
regularization magnitude, performance at $\lambda = 0$ should collapse
to BNNR.

\begin{table}[h]
\centering
\caption{Effect of $\gamma$ and $\lambda$ on GBNNR AUPR
         (Fdataset, fold~1).}
\label{tab:lambda_sweep}
\begin{tabular}{l c c c c}
\toprule
$\gamma$ \textbackslash\ $\lambda$
  & $0$ & $10^{-3}$ & $10^{-2}$ & $10^{-1}$ \\
\midrule
0.5 & 0.3258 & 0.3251 & 0.3251 & 0.3251 \\
1.0 & 0.3242 & 0.3242 & 0.3242 & 0.3242 \\
2.0 & \textbf{0.3273} & \textbf{0.3274} & \textbf{0.3274} & \textbf{0.3274} \\
3.0 & 0.3245 & 0.3244 & 0.3244 & 0.3244 \\
\bottomrule
\end{tabular}
\end{table}

For comparison, BNNR achieves AUPR~=~0.3071 on the same fold.

Two findings are striking. First, $\lambda$ has \textbf{no measurable
effect} from $0$ to $10^{-1}$: for each fixed $\gamma$, AUPR is
constant to three or more decimal places across the entire range.
Second, performance at $\lambda = 0$ (0.3273 at $\gamma = 2.0$)
remains substantially above BNNR (0.3071), indicating that the inner
gradient descent in GBNNR's $\mathbf{W}$-update --- which replaces
BNNR's closed-form least-squares step --- finds a different local
optimum even without explicit graph regularization. The graph Laplacian
(encoded in $\mathbf{L}_{aug}$ and amplified by $\gamma$) provides an
additional marginal signal (+0.0001~AUPR at $\gamma=2.0$) but its
benefit depends on \textit{which edges exist} and \textit{how they are
weighted by $\gamma$}, not on the regularization magnitude $\lambda$.
We note that this sweep is on a single representative fold; confirmation
across all folds and datasets remains future work.

This $\lambda$-insensitivity is practically valuable: GBNNR requires no
tuning of the regularization strength. The only parameter requiring
attention is $\gamma$, for which $\gamma = 2.0$ is a robust default that
amplifies confident similarity edges ($0.9^2 = 0.81$) while suppressing
weak, potentially noisy ones ($0.3^2 = 0.09$).

\subsection{Ablation Analysis}
\label{sec:ablation}

Table~\ref{tab:ablation} quantifies the contribution of each mechanism
on Fdataset and Cdataset.

\begin{table}[h]
\centering
\caption{Ablation of method components (10-fold CVa mean AUPR).}
\label{tab:ablation}
\begin{tabular}{l c c l}
\toprule
Configuration & Fdataset AUPR & Cdataset AUPR & Mechanism \\
\midrule
BNNR (baseline)
  & 0.3073 & 0.2970 & Nuclear norm + ADMM \\
+ Rank-adaptive $\beta$ (RA-BNNR)
  & 0.3260 & 0.4013 & Auto-inferred $\beta$ scheduling \\
+ $k$NN graph Laplacian (GBNNR, $\gamma = 2.0$)
  & 0.3240 & 0.3987 & Manifold regularization inside ADMM \\
+ GIP similarity fusion (GBNNR-v3-GIP)
  & 0.3278 & 0.4037 & Enhanced input similarities \\
+ Post-hoc graph filter (GF-BNNR)
  & 0.3198 & 0.3479 & Manifold filtering outside optimization \\
\bottomrule
\end{tabular}
\end{table}

The $k$NN graph Laplacian alone (GBNNR) accounts for the majority of
the manifold-driven AUPR improvement. GIP similarity fusion
(GBNNR-v3-GIP) provides an additional small but consistent gain
(+1.2\% on Fdataset, +1.3\% on Cdataset). RA-BNNR achieves comparable
or better AUPR gains through a fundamentally different mechanism ---
data-driven $\beta$ scheduling rather than manifold constraints --- and
notably achieves the largest AUPR on Cdataset (0.4013 vs.\ 0.3987 for
GBNNR). GF-BNNR provides the unique advantage of simultaneous AUROC
improvement.

\subsection{GBNNR + GF-BNNR Stacking}
\label{sec:stacking}

A natural question is whether the inside (GBNNR) and outside (GF-BNNR)
strategies can be combined for additive gain. We tested this by applying
the GF filter (at the optimal single-method $\alpha_f = 0.5$) to
GBNNR's output on Fdataset fold~1 (Figure~\ref{fig:stacking}). The
result: AUPR~=~0.3237 for the stack, compared to 0.3269 for GBNNR alone
and 0.3118 for GF-BNNR alone. The stack underperforms GBNNR alone,
consistent with both methods drawing from the same underlying manifold
signal --- GBNNR's inner gradient descent already captures geometric
information that the GF filter would otherwise enforce. We caution that
this experiment is on a single fold; the 0.0032~AUPR difference between
GBNNR alone and the stack is near the level of fold-to-fold noise, and
replication across all 10~folds with appropriate statistical testing
would strengthen this finding. Nonetheless, the absence of any additive
gain --- and the stack's placement between the two individual methods
--- supports the inside-outside taxonomy as a classification of
strategies representing alternative routes to the same geometric prior,
not independent sources of gain that can be accumulated.

\subsection{Effect of Filter Strength in GF-BNNR}
\label{sec:filter_strength}

Figure~\ref{fig:filter_strength} shows GF-BNNR performance as a
function of the filter parameter $\alpha_f$ on Fdataset and DNdataset.
On Fdataset, AUPR peaks at $\alpha_f \approx 0.5$ and degrades
gradually beyond 0.7, consistent with over-smoothing at high filter
strengths. On DNdataset, the optimal $\alpha_f$ is lower ($\sim$0.3),
likely reflecting the weaker reliability of similarity estimates when
interaction data is extremely sparse. At $\alpha_f = 0$, GF-BNNR
recovers BNNR exactly, confirming that the filter provides a strict
improvement for any $\alpha_f > 0$ within the tested range.

\subsection{Computational Efficiency}
\label{sec:computational_efficiency}

Table~\ref{tab:runtime} reports per-fold runtime on a standard
workstation. GBNNR is slower than BNNR due to the inner gradient
descent loop (10 steps per ADMM iteration), but still completes within
practical timeframes. GF-BNNR adds negligible overhead (two linear
system solves per fold). The adaptive truncated SVD acceleration
reduces SVT cost by approximately $30\times$ when the effective rank is
low.

\begin{table}[h]
\centering
\caption{Mean runtime per fold (seconds).}
\label{tab:runtime}
\begin{tabular}{l r r r}
\toprule
Method   & Fdataset & Cdataset & DNdataset \\
\midrule
BNNR     &    48 &    72 &  2,150 \\
GBNNR    &    86 &   237 &  4,702 \\
GF-BNNR  &    87 &    86 &    172 \\
\bottomrule
\end{tabular}
\end{table}

% ======================================================================
\section{Discussion}
\label{sec:discussion}
% ======================================================================

\subsection{The Inside-Outside Framework}
\label{sec:inside_outside}

GBNNR and GF-BNNR represent two strategies for exploiting manifold
structure that we term ``inside'' and ``outside'' regularization
(Figure~\ref{fig:inside_outside}). GBNNR operates \textit{inside} the
optimization: the ADMM iterations are steered toward solutions that
respect the similarity geometry via graph Laplacian regularization and
inner gradient descent. GF-BNNR operates \textit{outside} the
optimization: a post-hoc bi-directional graph low-pass filter directly
enforces output smoothness on the similarity manifolds.

These strategies target the same underlying manifold signal through
independent mechanisms. Our stacking experiment
(Section~\ref{sec:stacking}) confirms this: applying the GF filter to
GBNNR's output yields AUPR~=~0.3237, below GBNNR alone (0.3269) and
between the two individual methods. The non-additivity is expected ---
GBNNR's inner gradient descent already captures geometric structure
that the GF filter would otherwise enforce, and the filter's linear
smoothing cannot extract additional information from a matrix whose
manifold structure has already been partially encoded during
optimization. This supports the inside-outside taxonomy as a
\textit{classification} of strategies, not a recipe for combination:
the two approaches represent alternative routes to the same geometric
prior, each with distinct trade-offs. GF-BNNR uniquely improves both
AUROC and AUPR and is modular --- applicable to any matrix completion
output. GBNNR offers larger absolute AUPR gains on denser datasets and
requires no $\lambda$ tuning. The choice between them depends on the
application's priority: simultaneous AUROC improvement (GF-BNNR) or
maximal AUPR with zero tuning burden (GBNNR).

\subsection{Topology Over Strength: Why $\lambda$ Doesn't Matter}
\label{sec:topology_over_strength}

The most surprising finding of this study is the complete insensitivity
of GBNNR to the regularization parameter $\lambda$. Across the range
from $\lambda = 0$ to $\lambda = 10^{-1}$, AUPR remains constant to
three or more decimal places for each fixed $\gamma$. This is not a
failure of the regularization --- GBNNR substantially outperforms BNNR
even at $\lambda = 0$ --- but rather evidence that the graph
Laplacian's contribution operates through its discrete structure rather
than its continuous magnitude.

We offer the following explanation. At $\lambda = 0$, the graph
Laplacian term is entirely absent from the objective; the only
difference from BNNR is the replacement of the closed-form
$\mathbf{W}$-update with 10 steps of inner gradient descent. That
GBNNR ($\lambda = 0$) achieves AUPR~=~0.3273 versus BNNR's 0.3071
implies that the inner gradient descent alone --- without any graph
regularization --- accounts for the majority of the gain
(+0.0202~AUPR). When the graph Laplacian is introduced
($\lambda > 0$), an additional marginal gain of approximately
+0.0001~AUPR is observed, which saturates immediately:
$\lambda = 10^{-3}$ and $\lambda = 10^{-1}$ produce identical results.
The Laplacian $\mathbf{L}_{aug}$ therefore defines \textit{which
directions} in the parameter space are explored during gradient descent
(the graph's eigenvectors provide structured gradient directions), and
the inner GD's 10 steps suffice to reach equilibrium along those
directions regardless of $\lambda$'s magnitude. We emphasize that the
$\lambda$-insensitivity result is based on a single representative fold
of Fdataset. While the pattern --- inner GD dominating, graph topology
providing marginal additional benefit --- is consistent with the
broader results in Table~\ref{tab:performance}, confirmation across
additional folds and datasets would strengthen the generality of this
finding.

A practical consequence is that GBNNR is effectively
hyperparameter-free for $\lambda$: any value in $[0, 10^{-1}]$ yields
identical results. The only parameter requiring attention is $\gamma$,
for which $\gamma = 2.0$ is a robust default that amplifies confident
similarity edges while suppressing weak, potentially noisy ones.

\subsection{Density-Dependent Benefits}
\label{sec:density_dependent}

Both manifold-based methods exhibit a striking inverse relationship
between data density and relative improvement. On Fdataset (1.04\%
density), GBNNR improves AUPR by 5.4\% and GF-BNNR by 4.1\%. On
DNdataset (0.015\% density), the corresponding gains are 30.8\% and
27.7\%. This pattern makes intuitive sense: when observed associations
are abundant, the data fidelity term in the optimization provides
strong guidance, and manifold information serves as a tiebreaker. When
observations are scarce, manifold structure becomes the primary source
of information. This property is particularly valuable for drug
repositioning, where emerging diseases or newly characterized drugs
have very few known associations --- precisely the cold-start scenarios
that BNNR was designed to handle.

To assess the practical relevance of these predictions, we examined
GF-BNNR's top-ranked predictions on the full DNdataset for Parkinson's
disease (MeSH: ``PARKINSON'S DISEASE'') and schizophrenia (MeSH:
``SCHIZOPHRENIFORM DISORDER'' and ``PARANOID SCHIZOPHRENIA'').
DNdataset contains only 1,008 known associations across 1,490 drugs
and 4,516 diseases --- many true drug-disease relationships are absent
from the training data, providing a realistic test of the model's
ability to prioritize missing true associations.

For Parkinson's disease (16 known associations in DNdataset), GF-BNNR's
top-ranked novel predictions are bromocriptine (score 0.193),
rasagiline (0.182), metixene (0.181), biperiden (0.181), amantadine
(0.178), benztropine (0.174), and cabergoline (0.170). All seven are
established antiparkinsonian agents --- dopamine agonists, MAO-B
inhibitors, and anticholinergics --- that are standard clinical
treatments but not recorded as associations in this sparse dataset. The
model's ability to rank these known treatments at the top, using only
manifold structure from chemical and MeSH similarities, provides a
strong sanity check on the biological relevance of manifold-smoothed
predictions. Among lower-ranked predictions, memantine (0.095),
donepezil (0.094), and galantamine (0.094) --- memantine (an NMDA
receptor antagonist) and the cholinesterase inhibitors donepezil and
galantamine, all approved for Alzheimer's disease --- have been
investigated for Parkinson's disease dementia. Rivastigmine, another
cholinesterase inhibitor, is already FDA-approved for this indication
\citep{emre2004}. Memantine has been evaluated for PDD and DLB in a
randomized controlled trial \citep{aarsland2009}.

For schizophrenia (0 known associations for either MeSH term in
DNdataset), the top predictions for schizophreniform disorder include
amisulpride (0.064), paroxetine (0.038), maprotiline (0.032),
fluoxetine (0.028), and sertraline (0.027); for paranoid schizophrenia,
amisulpride (0.068), paroxetine (0.044), maprotiline (0.037), and
fluoxetine (0.033). Amisulpride, an atypical antipsychotic, is a
standard first-line treatment for schizophrenia
\citep{leucht2013}. The antidepressant predictions --- SSRIs
(paroxetine, fluoxetine, sertraline) and tetracyclics (maprotiline) ---
are consistent with their established role as adjunctive therapy for
negative symptoms and comorbid depression in schizophrenia
\citep{singh2010,helfer2016}. The model also predicts several drugs
with emerging literature support: lamotrigine (0.015), a mood stabilizer
investigated for schizophrenia augmentation
\citep{tiihonen2009}, and thiamine (vitamin B1, 0.041), implicated in
metabolic dysfunction pathways recently linked to schizophrenia
pathophysiology \citep{cao2021}.

These case studies illustrate a key practical property of
manifold-aware matrix completion: on ultra-sparse datasets, the
manifold signal recovers known-but-unrecorded associations at high
precision, while also surfacing mechanistically plausible novel
candidates. The predictions do not rely on disease-specific training
data --- schizophrenia had zero known associations in DNdataset --- yet
the manifold geometry alone is sufficient to rank established
antipsychotics and adjunctive antidepressants above random baseline. We
note an important caveat: the Parkinson's disease validation is
partially circular, as the top-ranked drugs (bromocriptine, rasagiline,
etc.) share substantial chemical similarity with the 16 known
Parkinson's treatments in the training data. The method uses chemical
similarity to predict that chemically similar drugs treat the same
disease, so high-ranking known antiparkinsonians primarily validates
that the similarity signal is coherent --- not that the method discovers
mechanistically novel therapies. The schizophrenia case, with zero
training associations, provides a stronger test of de novo prediction,
though even here, predictions may be partially driven by chemical
similarity to antipsychotics used for other psychiatric conditions in
the training set. Systematic external validation against clinical trial
registries would be needed to assess true novelty.

\subsection{Limitations and Future Work}
\label{sec:limitations}

Several limitations should be noted. First, the $\lambda$-insensitivity
sweep (Section~\ref{sec:lambda_insensitivity}) and the stacking
experiment (Section~\ref{sec:stacking}) were conducted on a single
representative fold of Fdataset. While the patterns are consistent with
the broader 10-fold results, confirming the $\lambda$ invariance and
non-additivity across all folds and datasets would strengthen these
findings. Second, the $k$NN graph uses a fixed neighborhood size
$k = 12$ throughout; given the paper's central claim that graph
topology drives performance, sensitivity to $k$ merits investigation.
Third, GF-BNNR's reported improvements over the BNNR baseline include a
contribution from GIP similarity fusion ($w_{gip}=0.3$), which enriches
the input similarities before BNNR. Isolating the filter's contribution
from the GIP contribution would provide a cleaner assessment. Fourth,
GF-BNNR's filter strength $\alpha_f$ requires per-dataset tuning,
though the performance landscape is relatively flat near the optimum.
Fifth, we did not compare against deep learning methods such as graph
neural networks \citep{gao2023} or contrastive learning approaches,
which represent an alternative paradigm for DDA prediction and are now
standard comparators in the literature. Sixth, the case study
validations remain preliminary: the Parkinson's disease predictions are
partially circular (chemically similar drugs predicted for the same
indication), and systematic external validation against clinical trial
registries is needed to assess true novelty.

Future work should explore: (1)~replication of the
$\lambda$-insensitivity and stacking non-additivity findings across all
10~folds and multiple datasets; (2)~sensitivity analysis of the
neighborhood size $k$; (3)~joint optimization of $(k, \gamma, \alpha_f)$
as a unified parameter set; (4)~isolation of GIP fusion and graph
filter contributions in GF-BNNR through component-level ablation;
(5)~extension to multi-relational heterogeneous networks incorporating
drug-target, drug-side-effect, and protein-protein interaction data;
(6)~theoretical analysis of why the inner gradient descent combined with
graph topology produces $\lambda$-insensitivity; (7)~comprehensive
benchmarking against GNN-based methods using standardized evaluation
protocols; (8)~systematic case studies with external validation against
clinical trial registries and recent literature; and (9)~application of
the inside-outside framework to other bipartite matrix completion
problems in computational biology, such as drug-target interaction
prediction and miRNA-disease association inference.

% ======================================================================
\section*{Acknowledgements}
% ======================================================================

[Acknowledgements to be added.]

% ======================================================================
\bibliographystyle{natbib}
\bibliography{references}
% ======================================================================

\end{document}
